---
title: "STAT 205 HW4"
author: "HONGLEI REN"
date: "03/04/2020"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list=ls())
library(knitr)
library(readr)
library(broom)
library(tidyverse)
library(R2jags)
library(epiR)
library(bayesplot)
library(coda)
library(robustHD)
library(patchwork)
```

# EX 9.21

### (a) Data analysis
```{r}
  FEVdata <- read.table(file="FullFEVdata.txt",header=T, sep="") 
  attach(FEVdata)
  pairs(~Age+Hgt+Male+Smoke+FEV,labels=c("Age","Height","Male","Smoke","FEV"),data=FEVdata)
```
From the figure above, we can see that Age and Height are postively correlated and FEV increases as either of them increase. Smoke-adolescents in this study range from 7 to 18 years old, and have higher FEV.

### (b) Prior, Predictors and Model Diagnosis
#### 1) Predictor selection
I chose indicators of smoke, age, height, interactions between age and smoke as the predictors, which are suggested by data analysis.

```{r}
  Age.st=standardize(Age)
  FEV.st=standardize(FEV)
  X.mat=model.matrix(~ as.factor(Smoke) + Age.st  + as.factor(Smoke)*Age.st)
```
#### 2) Prior Construction
Here, I apply the BCJ independent priors for $\beta$ and $\tau$. According to the given prior knowledge from Dr. David. We have:

$$\widetilde{X} = \begin{bmatrix}
\widetilde{x_1}'\\ 
\widetilde{x_2}'\\ 
\widetilde{x_3}'\\ 
\widetilde{x_4}'
\end{bmatrix}=  \begin{bmatrix}
1 & 1 & 18 & 18  \\ 
1 & 0  & 16 & 0  \\ 
1 & 1 & 13 & 13 \\ 
1 & 0 & 12 & 0
\end{bmatrix}$$

$$
\widetilde{X}\beta = \widetilde{m} = \begin{bmatrix}
\widetilde{m_1}'\\ 
\widetilde{m_2}'\\ 
\widetilde{m_3}'\\ 
\widetilde{m_4}'
\end{bmatrix}=  N(\begin{bmatrix}
4.0\\ 
4.2\\ 
3.4\\ 
2.7
\end{bmatrix}, \begin{bmatrix}
0.118 & 0 & 0 & 0 \\ 
0 & 0.118 & 0 & 0 \\ 
0 & 0 & 0.067 & 0 \\ 
0 & 0 & 0 & 0.118
\end{bmatrix})
$$

Therefore, we got the prior for $\beta$ is:
$$
\beta =\widetilde{X}^{-1} \widetilde{m} = \widetilde{X}^{-1}  \begin{bmatrix}
\widetilde{m_1}'\\ 
\widetilde{m_2}'\\ 
\widetilde{m_3}'\\ 
\widetilde{m_4}'
\end{bmatrix}=  N(\begin{bmatrix}
-1.80\\
3.64\\
0.375\\
-0.255\\
\end{bmatrix}, \begin{bmatrix}
2.9500  & -2.9500  & -0.2065  &  0.2065\\
   -2.9500  &  4.6160   & 0.2065  & -0.3161\\
   -0.2065 &   0.2065 &   0.0147  & -0.0147\\
    0.2065 &  -0.3161 &  -0.0147  &  0.0221\\
\end{bmatrix})
$$

For $\tau$, I assume $gamma(0.001, 0.001)$ as prior.

```{r message=F}
Ytilde <- c(4., 4.2, 3.4, 2.7) # Specify prior mean vector
D <- diag(c(0.118, 0.118, 0.067, 0.118)) 
Xtilde <- matrix(c(1, 1, 1, 1, 1, 0, 1, 0, 18, 16, 13, 12, 18, 0, 13, 0),4,4,byrow=F)
Xtildeinv <- solve(Xtilde) # Invert Xtilde
# Get prior mean vector and precision matrix for beta
beta0 <- c(t(Xtildeinv %*% Ytilde))
C0 <- Xtildeinv %*% D %*% t(Xtildeinv)
C0inv <- solve(C0)

jags.data=list(
  Y=FEV,
  Xmat=X.mat,
  r=dim(X.mat)[2],
  n=dim(X.mat)[1],
  beta0=beta0,
  C0inv=C0inv,
  a=0.001, b=0.001## diffuse prior
)

model_reg <- "model{
for(i in 1:n){
  Y[i] ~ dnorm(mu[i], tau)
  mu[i] <- beta[1] + beta[2]*Xmat[i,2] + beta[3]*Xmat[i,3] + beta[4]*Xmat[i,4]
}
beta[1:r] ~ dmnorm(beta0[1:r], C0inv[1:r, 1:r]) # g prior 
tau ~ dgamma(a, b)

for(i in 1:19){
meanFEVs[i] <- beta[1] + beta[2]+ (beta[3]+beta[4])* i 
meanFEVns[i] <- beta[1] + beta[3]* i
}
FEV20s ~ dnorm(mu20s,tau)
FEV20ns ~ dnorm(mu20ns,tau)
mu20s <- beta[1] + beta[2]+ (beta[3]+beta[4])* 20
mu20ns <- beta[1] + beta[3]*20 
}"

jags.param=c("beta","tau","meanFEVns[7]","meanFEVns[8]","meanFEVns[9]","meanFEVns[10]","meanFEVns[11]", "meanFEVns[12]", "meanFEVns[13]", "meanFEVns[14]", "meanFEVns[15]", "meanFEVns[16]", "meanFEVns[17]", "meanFEVns[18]", "meanFEVns[19]", "meanFEVs[7]","meanFEVs[8]","meanFEVs[9]","meanFEVs[10]","meanFEVs[11]", "meanFEVs[12]", "meanFEVs[13]", "meanFEVs[14]", "meanFEVs[15]", "meanFEVs[16]", "meanFEVs[17]", "meanFEVs[18]", "meanFEVs[19]", "FEV20s","FEV20ns") 
```
```{r echo=T, results='hide'}
jags.fit <- jags(data=jags.data, parameters.to.save = jags.param,
model.file=textConnection(model_reg), n.iter=20000, n.chains=1, n.burnin=5000, n.thin=1, DIC=T, digits=6)
```
#### 3) Convergence and model diagnosis

```{r message=F}
jags.mcmc = as.mcmc(jags.fit) 
mcmc_trace(jags.mcmc, pars = c("beta[1]", "tau"))
```


From `beta[1]` and `tau` in the fitted MCMC chain, we can see that the parameters are converged because it is well-mixed, no zig-zag, and the model also shows an white-noise like error term `tau`, which suggests that the covariates are independent.

### c) Posterior inferences
The inferred posterior paramters are listed below:
```{r}
jags.fit
```

```{r}
library(bayesplot)
par(mfrow=c(1,2))

mcmc_dens(jags.mcmc, pars=c("meanFEVns[18]", "meanFEVs[18]")) 
mcmc_dens(jags.mcmc, pars=c("FEV20ns", "FEV20s")) 
```

From the plot above, we can see that non-smokers are higher in both posterior FEV of 18-yrs-old group and predictive FEV of 20-yrs-old population.

### d)  normal FEV ranges for several different types of adolescents

The FEV ranges of non-smokers and smokers for different ages are shown below:
```{r message=F}
df_mcmc = data.frame(jags.mcmc[[1]])  %>% 
  select(starts_with("meanFEV"))
colnames(df_mcmc) <- gsub('meanFEVs.', 'smoker age ', colnames(df_mcmc), fixed=TRUE)
colnames(df_mcmc) <- gsub('meanFEVns.', 'non smoker age ', colnames(df_mcmc), fixed=TRUE)
df_mcmc %>% 
  lapply(quantile, prob = c(0.05, 0.95), na.rm = TRUE) %>% 
  data.frame() %>% 
  t() %>% 
  kable()
```
For easy reading the result, I visulized the ranges as follows:
```{r message=F}
color_scheme_set("blue")
par(mfrow=c(1,2))
p1 = mcmc_intervals(jags.mcmc, pars = c("meanFEVns[7]","meanFEVns[8]","meanFEVns[9]","meanFEVns[10]","meanFEVns[11]", "meanFEVns[12]", "meanFEVns[13]", "meanFEVns[14]", "meanFEVns[15]", "meanFEVns[16]", "meanFEVns[17]", "meanFEVns[18]", "meanFEVns[19]"))
p2 = mcmc_intervals(jags.mcmc, pars = c("meanFEVs[7]","meanFEVs[8]","meanFEVs[9]","meanFEVs[10]","meanFEVs[11]", "meanFEVs[12]", "meanFEVs[13]", "meanFEVs[14]", "meanFEVs[15]", "meanFEVs[16]", "meanFEVs[17]", "meanFEVs[18]", "meanFEVs[19]"))
p1 + p2
```

### e) Sensitivity Analysis
By changing the `a = b = 0.001` to `a = b = 0.1`, we perfromed same pipeline, and got the ranges of FEV as follows, which do not show significant changes compared with previous one.
```{r}
jags.data=list(
  Y=FEV,
  Xmat=X.mat,
  r=dim(X.mat)[2],
  n=dim(X.mat)[1],
  beta0=beta0,
  C0inv=C0inv,
  a=0.1, b=0.1## diffuse prior
)
jags.fit <- jags(data=jags.data, parameters.to.save = jags.param,
model.file=textConnection(model_reg), n.iter=20000, n.chains=1, n.burnin=5000, n.thin=1, DIC=T, digits=6)

jags.mcmc = as.mcmc(jags.fit) 
par(mfrow=c(1,2))
p1 = mcmc_intervals(jags.mcmc, pars = c("meanFEVns[7]","meanFEVns[8]","meanFEVns[9]","meanFEVns[10]","meanFEVns[11]", "meanFEVns[12]", "meanFEVns[13]", "meanFEVns[14]", "meanFEVns[15]", "meanFEVns[16]", "meanFEVns[17]", "meanFEVns[18]", "meanFEVns[19]"))
p2 = mcmc_intervals(jags.mcmc, pars = c("meanFEVs[7]","meanFEVs[8]","meanFEVs[9]","meanFEVs[10]","meanFEVs[11]", "meanFEVs[12]", "meanFEVs[13]", "meanFEVs[14]", "meanFEVs[15]", "meanFEVs[16]", "meanFEVs[17]", "meanFEVs[18]", "meanFEVs[19]"))
p1 + p2
```

### f) Summary the Analysis

Above, we constructed a linear model by using age, smoke, and their interaction as predictor. For the priors, we used BCJ priors for $\beta$(coefficients), and Gamma prior for $\tau$ (error term). From the inferred and predictive FEVs, we can see that smokers generally have lower FEVs than non-smokers, and as the age increasing, FEV is increasing as well.

# EX 9.22
```{r message=F}
coleman = read_csv(file="Coleman.csv")  
attach(coleman)
jags.data=list(
  y=y,
  x=x,
  n=dim(coleman)[1],
  b=0.01, c=0.001## diffuse prior
)

model_reg <- 
"model{ 
  for(i in 1:n){
      y[i] ~ dnorm(mu[i], tau)
      mu[i] <- beta[1] + beta[2]* x[i]
  }
  for(j in 1:2){
    beta[j] ~ dmnorm(0, b)
  }
  
  tau ~ dgamma(c, c)
  pred = beta[1] + beta[2]*(-16.04)
}"

jags.param=c("tau", "beta", "pred")
```
```{r echo=T, results='hide'}
jags.fit <- jags(data=jags.data, parameters.to.save = jags.param,
                 model.file=textConnection(model_reg), n.iter=20000, n.chains=1,
                 n.burnin=5000,n.thin=1, DIC=T, digits=6)
```

```{r echo = FALSE, message =F}
coleman %>% 
  ggplot() + 
  aes(x = x, y = y) +
  geom_point(color = "blue") +
  geom_smooth(method = "lm")
```

From the fitted result above, `x` and `y` are positively correlated with `r cor(x, y)` Pearson correlation.

```{r}
jags.fit
jags.mcmc = as.mcmc(jags.fit) 
mcmc_trace(jags.mcmc, pars = c("beta[2]", "tau"))
```

#### From the inference above, the  predictive value of $x = −16.04$ is $24.161$, with 95% credible interval (21.693, 26.535).

```{r}
mcmc_dens(jags.mcmc, pars = c("pred"))
```

#### I am not surprised that higher socioeconomic status were positively associated with higher test scores, because they can receive better education resource(environment, teaching method, etc.)



